Measuring contrast processing in the visual system using SSVEP

### 2024-04-08 review.qmd
Datathief to the rescue! Including parameters from Afsari et al 2014

A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011]. The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.

Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction. For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof. This corresponds closely to similar functions reported by invasive studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982]. However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.

@@

fig, axs = plt.subplots(1,3, figsize=(8, 3))

@@

The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains a slight roll-off at high input contrasts. This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity. Power at other harmonics increases, and the total power generated by the input is monotonic. This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].

@@

fig, axs = plt.subplots(1,3, figsize=(8, 3))

inputModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]

axs[0].plot(support_S, inputModulation,'k')
axs[0].plot(support_S, maskModulation,'r')
axs[0].set_title('Input modulations', fontsize=10)
axs[0].grid()

combinedSignal=simpleTransducerFunction(inputModulation, maskModulation, c50,RMax)
FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
axs[1].plot(support_S, combinedSignal,'k')
axs[1].set_title('Combined signal after transducer', fontsize=10)
axs[1].grid()

@@

This possibility was recognised in the early days of the technique [@Regan1988]. Although characterising the complete set of computations along entire processing pathway is problematic, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data. ... MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much. @Candy2001; @Busse2009; @Tsai2012; @Baker2017

@@

Because SSVEP responses at detection threshold are very small, estimating threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep). This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].

@@

Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979]. To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990]. The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity. In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].

This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979]. In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months. Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson1983]. At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction. However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].

@@

Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022]. @Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls. This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker). Interestingly this study replicated its key findings in a *drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.

@@

# Assuming Parkinson's parameters are similar to those given for epilepsy/autism,
# but with specifics for Parkinson's. We'll use the epilepsy parameters as a placeholder.

# Note: Replace epilepsyparams with Parkinson's specific parameters as necessary.
respParkControls = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
                      (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]

respParkPatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
                      (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='Parkinson\'s patients')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='Control participants')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 7, 2))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
# To save the figure
# plt.savefig('Figures/clinicalplot.pdf')
plt.show()

>>>>

A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011]. The SSVEP has proven to be invaluable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.

Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction. For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof. This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982]. However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.

@@

#Make a figure with three subplots
fig, axs = plt.subplots(1,3, figsize=(8, 3))

@@

The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts. This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity. Power at other harmonics increases, and the total power generated by the input is monotonic. This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].

@@

# Make a figure with three subplots
fig, axs = plt.subplots(1,3, figsize=(8, 3))

# Compute waveforms for the 'probe' and 'mask' inputs at the frequencies defined in inputFrequency_Hz
probeModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]

axs[0].plot(support_S, probeModulation,'k')
axs[0].plot(support_S, maskModulation,'r')
axs[0].set_title('Input modulations', fontsize=10)
axs[0].grid()

# pass both the probe and the mask through the transducer
combinedSignal=simpleTransducerFunction(probeModulation, maskModulation, c50,RMax)

# Compute the FT of the combined signals
FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
axs[1].plot(support_S, combinedSignal,'k')
axs[1].set_title('Combined signal after transducer', fontsize=10)
axs[1].grid()

@@

This possibility was recognised in the early days of the technique [@Regan1988]. Although characterising the complete set of computations along entire processing pathway is problematic, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data. ... MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much. @Candy2001; @Busse2009; @Tsai2012; @Baker2017 *** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS 'FAST' VEPs ***

@@

Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep). This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].

@@

A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels. This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979]. To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990]. The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity. In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].

This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979]. In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months. Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983]. At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction. However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].

@@

Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022]. @Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls. This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker). Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.

@@

# PD parameters for Drosophila from Afsari et al HMG 2014
PDparams2F1 = np.array([
    [22, 0.7, 2, 1],  # controls
    [24, 0.4, 2, 1]])   # hLRRK2

PDparams1F1 = np.array([
    [104, 0.76, 2, 1],  # controls
    [170, 0.382, 2, 1]])   # hLRRK2

# Note: Here we plot the 2F1 curves which, we think, most closely correspond to the more central neuronal (as opposed to the photoreceptor-derived) signals
respParkControls = PDparams2F1[0,0] * ((finecontrasts/100)**PDparams2F1[0,2] /
                      (PDparams2F1[0,1]**PDparams2F1[0,2] + (finecontrasts/100)**PDparams2F1[0,2])) + PDparams2F1[0,3]

respParkPatients = PDparams2F1[1,0] * ((finecontrasts/100)**PDparams2F1[1,2] /
                      (PDparams2F1[1,1]**PDparams2F1[1,2] + (finecontrasts/100)**PDparams2F1[1,2])) + PDparams2F1[1,3]

plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='hLRRK2 transgenic flies')
plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='TH+ (control) flies')
plt.xticks(ticklocsx, ticklabelsx)
plt.yticks(np.arange(0, 7, 2))
plt.xlabel('Target contrast (%)')
plt.ylabel('SSVEP response')
plt.title("(d) Parkinson's")
plt.legend()
# To save the figure
# plt.savefig('Figures/clinicalplot.pdf')
plt.show()

### 2024-04-08 review.tex
Datathief to the rescue! Including parameters from Afsari et al 2014

A significant body of research into contrast processing is concerned
with how these normalization mechanisms depend on colour (Chen \emph{et
al.}, 2000), orientation (Foley, 1994), eye of origin (Legge, 1979;
Baker \emph{et al.}, 2007), spatial and temporal frequency (Meese \&
Baker, 2009), location (Polat \& Sagi, 1993; Tadin \emph{et al.}, 2003;
Petrov \emph{et al.}, 2005), age (Betts \emph{et al.}, 2005), and the
presence of neurological disorders (Porciatti \emph{et al.}, 2000; Tsai
\emph{et al.}, 2011). The SSVEP has proven to be invaliable in this
research because it provides an objective readout of contrast
representation at different stages of the visual system, and allows us
to `tag' the probe and background at separate frequencies.

Because it provides a direct read-out of neural population activity, the
SSVEP signal can reveal key features of neural signal transduction. For
example, by varying the peak stimulus contrast parametrically, a
`contrast vs response' function can be measured - where the `response'
is typically defined as the amplitude of the SSVEP frequency component
at the stimulus frequency or a low multiple thereof. This corresponds
closely to similar functions reported by invasive studies measuring
single unit activity or local field potentials in the cortex (Shapley \&
Victor, 1980; Morrone \emph{et al.}, 1982). However the SSVEP has the
advantage that it is non-invasive, and so can be measured in awake,
behaving human participants.

@@

The hyperbolic ratio function is monotonic but the CRF resulting from
measuring the amplitude of the 2F component contains a slight roll-off
at high input contrasts. This results from the distortion of the input
sine waves at high contrast due to a combination of the full-wave
rectification and saturating non-linearity. Power at other harmonics
increases, and the total power generated by the input is monotonic. This
roll-off is often seen in experimental data and has been referred to as
`supersaturation' (Tyler \& Apkarian, 1985; Peirce, 2007).

@@

This possibility was recognised in the early days of the technique
(Regan \& Regan, 1988). Although characterising the complete set of
computations along entire processing pathway is problematic, careful
parametric variation of the input stimuli does allow us to fit models of
early visual gain control using SSVEP data. \ldots{} MORE HERE -
Although this idea was suggested as early as 1988 (probably earlier) not
many people seem to have done it much. Candy \emph{et al.} (2001); Busse
\emph{et al.} (2009); Tsai \emph{et al.} (2012); Baker \& Wade (2017)

@@

Because SSVEP responses at detection threshold are very small,
estimating threshold is achieved by measuring the contrast response
function at relatively high levels, and extrapolating back along the
function (either contrast vs response measured at a constant spatial
frequency or spatial frequency vs response at a constant contrast level)
to estimate its intercept with the x-axis (see
Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
approximately with psychophysically measured detection thresholds
(Norcia \emph{et al.}, 1986).

@@

Measuring the SSVEP at multiple levels was made faster by the
development of the `sweep VEP' paradigm in which the stimulus changed
its contrast, spatial frequency or some other property, throughout a
trial (Tyler \emph{et al.}, 1979). To avoid hysteresis effects, the
sweep is sometimes conducted both up and down in the same experiment
Norcia \& Tyler (1985\emph{a}). The sweep VEP (really, a sweep `SSVEP')
technique is now commonly-used to obtain a rapid and objective
measurement of visual acuity. In particular, because of its relative
speed and simplicity, this technique has now become a standard for
conducting tests of visual acuity in very young subjects or where
behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).

This approach has revealed much about the development of visual
abilities in infants (Harris \emph{et al.}, 1976; Atkinson \emph{et
al.}, 1979; Braddick \emph{et al.}, 1986). In general, SSVEP
measurements of infant vision have revealed that contrast sensitivity
for both achromatic and chromatic contrast as well as stereoscopic depth
perception develops earlier than had been supposed previously based on
behavioural readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler,
1985\emph{b}) with both chromatic and achromatic contrast detection
reaching near-adult levels by around six months. Spatial acuity as
measured by SSVEP reaches adult levels more slowly, but near-adult
levels are recorded around 1 year (Norcia \& Tyler, 1985\emph{b}; Hamer
\emph{et al.}, 1989) compared to around 6-7 years with behavioural
measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et al.}, 1999). At
least some of this difference is likely due to the relative objectivity
and high SNR of the SSVEP technique compared to other methods such as
preferential looking, which require careful measurement of the infant's
gaze direction. However it should be noted that other groups have
reported electrophysiological correlates of visual acuity that more
closely match the behavioural measures (De Vries-Khoe \& Spekreijse,
1982).

@@

Epilepsy is a neurological condition in which patients experience
seizures - episodes of uncontrolled neural activity that can cause
unconsciousness, involuntary movements and convulsions, and atypical
sensory experiences. Porciatti \emph{et al.} (2000) showed that
individuals with photosensitive epilepsy generate larger steady-state
signals in response to flickering visual stimuli, and their contrast
response functions saturate less than those of healthy controls. This is
consistent with the idea that epilepsy involves a cortical
hyperexciteability that makes seizures more likely. It is also the case
for individuals with idiopathic generalised epilepsy (Tsai \emph{et
al.}, 2011), a subtype of epilepsy that has a less obvious link to
vision. The differences apply across the whole contrast-response
function, and so resemble a response gain effect, which might be due to
reduced inhibition from neighbouring neurons. Differences in SSVEP
amplitudes have also been reported in individuals with migraine (Shibata
\emph{et al.}, 2008), a condition also associated with cortical
hyperexciteability.

@@

Autism is a condition often associated with differences in vision
(Simmons \emph{et al.}, 2009) and other senses (MacLennan \emph{et al.},
2022). Pei \emph{et al.} (2014) used a sweep-VEP method with
counterphase flickering stimuli, and found weaker responses in autistic
children at spatial frequencies around 8c/deg, compared with age-matched
controls. This was subsequently replicated in a further pediatric sample
by Vilidaite \emph{et al.} (2018) (see Figure~\ref{fig-plotclinical}c),
who additionally found weaker responses in autistic adults at the second
harmonic (using on/off flicker). Interestingly this study replicated its
key findings in a \emph{drosophila} genetic model of autism (\emph{Nhe3}
mutations), illustrating the translational potential of the steady-state
approach, as well as identifying a possible biomarker for autism.

@@

\includegraphics{review_files/figure-pdf/cell-8-output-1.pdf}

@@

\includegraphics{Figures/clinicalplot.pdf}

@@

\caption{\label{fig-plotclinical}caption.}

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Albrecht1982}
Albrecht DG \& Hamilton DB (1982).
\href{https://doi.org/10.1152/jn.1982.48.1.217}{Striate cortex of monkey
and cat: Contrast response function}. \emph{J Neurophysiol} \textbf{48,}
217--237.

>>>>

A significant body of research into contrast processing is concerned
with how these normalization mechanisms depend on colour (Chen \emph{et
al.}, 2000), orientation (Foley, 1994), eye of origin (Legge, 1979;
Baker \emph{et al.}, 2007), spatial and temporal frequency (Meese \&
Baker, 2009), location (Polat \& Sagi, 1993; Tadin \emph{et al.}, 2003;
Petrov \emph{et al.}, 2005), age (Betts \emph{et al.}, 2005), and the
presence of neurological disorders (Porciatti \emph{et al.}, 2000; Tsai
\emph{et al.}, 2011). The SSVEP has proven to be invaluable in this
research because it provides an objective readout of contrast
representation at different stages of the visual system, and allows us
to `tag' the probe and background at separate frequencies.

Because it provides a direct read-out of neural population activity, the
SSVEP signal can reveal key features of neural signal transduction. For
example, by varying the peak stimulus contrast parametrically, a
`contrast vs response' function can be measured - where the `response'
is typically defined as the amplitude of the SSVEP frequency component
at the stimulus frequency or a low multiple thereof. This corresponds
closely to similar functions reported by studies measuring single unit
activity or local field potentials in the cortex (Shapley \& Victor,
1980; Morrone \emph{et al.}, 1982). However the SSVEP has the advantage
that it is non-invasive, and so can be measured in awake, behaving human
participants.

@@

The hyperbolic ratio function is monotonic but the CRF resulting from
measuring the amplitude of the second harmonic (2F) component contains a
slight roll-off at high input contrasts. This results from the
distortion of the input sine waves at high contrast due to a combination
of the full-wave rectification and saturating non-linearity. Power at
other harmonics increases, and the total power generated by the input is
monotonic. This roll-off is often seen in experimental data and has been
referred to as `supersaturation' (Tyler \& Apkarian, 1985; Peirce,
2007).

@@

This possibility was recognised in the early days of the technique
(Regan \& Regan, 1988). Although characterising the complete set of
computations along entire processing pathway is problematic, careful
parametric variation of the input stimuli does allow us to fit models of
early visual gain control using SSVEP data. \ldots{} MORE HERE -
Although this idea was suggested as early as 1988 (probably earlier) not
many people seem to have done it much. Candy \emph{et al.} (2001); Busse
\emph{et al.} (2009); Tsai \emph{et al.} (2012); Baker \& Wade (2017)
*** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS `FAST' VEPs
***

@@

Because SSVEP responses at detection threshold are very small,
estimating a threshold is achieved by measuring the contrast response
function at relatively high levels, and extrapolating back along the
function (either contrast vs response measured at a constant spatial
frequency or spatial frequency vs response at a constant contrast level)
to estimate its intercept with the x-axis (see
Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
approximately with psychophysically measured detection thresholds
(Norcia \emph{et al.}, 1986).

@@

A robust estimate of the threshold therefore requires the measurement of
the SSVEP amplitude at many different super-threshold contrasr levels.
This was made faster by the development of the `sweep VEP' paradigm in
which the stimulus changed its contrast, spatial frequency or some other
property, throughout a trial (Tyler \emph{et al.}, 1979). To avoid
hysteresis effects, the sweep is sometimes conducted both up and down in
the same experiment Norcia \& Tyler (1985\emph{a}). The sweep VEP
(really, a sweep `SSVEP') technique is now commonly-used to obtain a
rapid and objective measurement of visual acuity. In particular, because
of its relative speed and simplicity, this technique has now become a
standard for conducting tests of visual acuity in very young subjects or
where behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et
al.}, 2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).

This approach has revealed much about the development of visual
abilities in infants (Harris \emph{et al.}, 1976; Atkinson \emph{et
al.}, 1979; Braddick \emph{et al.}, 1986). In general, SSVEP
measurements of infant vision have revealed that contrast sensitivity
for both achromatic and chromatic contrast as well as stereoscopic depth
perception develops earlier than had been supposed previously based on
behavioural readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler,
1985\emph{b}) with both chromatic and achromatic contrast detection
reaching near-adult levels by around six months. Spatial acuity as
measured by SSVEP reaches adult levels more slowly, but near-adult
levels are recorded around one year (Norcia \& Tyler, 1985\emph{b};
Hamer \emph{et al.}, 1989) compared to around six to seven years with
behavioural measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et
al.}, 1999). At least some of this difference is likely due to the
relative objectivity and high SNR of the SSVEP technique compared to
other methods such as preferential looking, which require careful
measurement of the infant's gaze direction. However it should be noted
that other groups have reported electrophysiological correlates of
visual acuity that more closely match the behavioural measures (De
Vries-Khoe \& Spekreijse, 1982).

@@

Epilepsy is a neurological condition in which patients experience
seizures - episodes of uncontrolled neural activity that can cause
unconsciousness, involuntary movements and convulsions, and atypical
sensory experiences. Porciatti \emph{et al.} (2000) showed that
individuals with photosensitive epilepsy generate larger steady-state
signals in response to flickering visual stimuli, and their contrast
response functions saturate less than those of healthy controls. This is
consistent with the idea that epilepsy involves a cortical
hyperexciteability that makes seizures more likely. It is also the case
for individuals with idiopathic generalised epilepsy (Tsai \emph{et
al.}, 2011), a subtype of epilepsy that has a less obvious link to
vision. The differences apply across the whole contrast-response
function, and so resemble a response gain effect (see
Figure~\ref{fig-plotclinical}a), which might be due to reduced
inhibition from neighbouring neurons. Differences in SSVEP amplitudes
have also been reported in individuals with migraine (Shibata \emph{et
al.}, 2008), a condition also associated with cortical
hyperexciteability.

@@

Autism is a condition often associated with differences in vision
(Simmons \emph{et al.}, 2009) and other senses (MacLennan \emph{et al.},
2022). Pei \emph{et al.} (2014) used a sweep-VEP method with
counterphase flickering stimuli, and found weaker responses in autistic
children at spatial frequencies around 8c/deg, compared with age-matched
controls. This was subsequently replicated in a further pediatric sample
by Vilidaite \emph{et al.} (2018) (see Figure~\ref{fig-plotclinical}c),
who additionally found weaker responses in autistic adults at the second
harmonic (using on/off flicker). Interestingly this study replicated its
key findings in a \emph{Drosophila} genetic model of autism (\emph{Nhe3}
mutations), illustrating the translational potential of the steady-state
approach, as well as identifying a possible biomarker for autism.

@@

Alex - it would be good to add some CRFs from Afsari \emph{et al.}
(2014) to the figure. Is it possible to extract either the fitted
parameter values or the raw data for some control flies and some
mutants? Whatever you think is most compelling.

@@

\includegraphics{review_files/figure-pdf/fig-plotclinical-output-1.pdf}

@@

\caption{\label{fig-plotclinical}Example contrast response functions for
different clinical conditions. Panel (a) shows modelled contrast
response functions for epilepsy patients (blue) and control participants
(black), based on the data of Tsai et al.~(2011). Panel (b) shows
functions for the amblyopic (red) and fellow (black) eyes of adults with
amblyopia, based on the data of Baker et al.~(2015). Panel (c) shows
functions for children with (green) and without (black) a diagnosis of
autism, based on the data of Vilidaite et al.~(2018). Panel (d)}

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Afsari2014}
Afsari F, Christensen KV, Smith GP, Hentzer M, Nippe OM, Elliott CJH \&
Wade AR (2014). \href{https://doi.org/10.1093/hmg/ddu159}{Abnormal
visual gain control in a {Parkinson}'s disease model}. \emph{Human
Molecular Genetics} \textbf{23,} 4465--4478.

\bibitem[\citeproctext]{ref-Albrecht1982}
Albrecht DG \& Hamilton DB (1982).
\href{https://doi.org/10.1152/jn.1982.48.1.217}{Striate cortex of monkey
and cat: Contrast response function}. \emph{J Neurophysiol} \textbf{48,}
217--237.
